### 2015-06-18 chapters/ipa_meets_unicode.tex
update homoglyphs in Unicode

  \item The uses of the apostrophe has led to long discussions on the Unicode
     Standard email list. An English keyboard inputs <{\fontspec{Monaco}'}> \textsc{apostrophe}
     at \uni{0027}, although the preferred Unicode apostrophe is the <\ ' >
     \textsc{right single quotation mark} at \uni{2019}.\footnote{Note that many word 
     processors (like Microsoft Word) by default will replace straight quotes by 
     curly quotes, depending on the whitespace around it.} However, the glottal stop/glottalization/ejective marker is yet 
     another completely different character <{\large ʼ}>, the 
     \textsc{modifier letter apostrophe} at \uni{02BC}, which unfortunately looks 
     mostly extremely similar to \uni{2019}. 
  \item Another problem is the <ˁ> \textsc{modifier letter reversed
     glottal stop} at \uni{02C1} vs.\@ the <ˤ> \textsc{modifier
     letter small reversed glottal stop} at \uni{02E4}. Both 
     appear in various resources representing phonetic data online. This is
     thus a clear example for which the Unicode Standard does not solve the
     linguistic standardization problem.
  \item even linguists are unlikely to distinguish between the <ə>
     \textsc{latin small letter schwa} at code point \uni{0259} and <ǝ>
     \textsc{latin small letter turned e} at \uni{01DD}.
  \item  The alveolar click <ǃ> at \uni{01C3} is of course often simply
     typed as <!> \textsc{exclamation mark} at \uni{0021}.\footnote{In the Unicode
     Standard the <ǃ> at \unif{01C3} is labeled \textsc{latin letter retroflex
     click}, but in IPA that glyph is used for an alveolar or postalveolar
     click (not retroflex). This naming is probably best seen as an 
     error in the Unicode Standard.}

@@

\begin{itemize}
  \item[] {\fontspec{Times New Roman}Times new Roman: <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] {\small \fontspec{CharisSIL}CharisSIL: <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] {\footnotesize \fontspec{Monaco}Monaco: <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] {\fontspec{DoulosSIL}DoulosSIL: <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] Linux Libertine: <t̥ʰ͡s> <t̥ʰ͜s>
\end{itemize}

Tie bars are a special type of character in the sense that they do not belong to
a segment, but bind two graphemes together. This actually turns out to be rather
different from the Unicode conceptions, see
Section~\ref{pitfall-different-notions-of-diacritics}. According to the spirit
of the IPA, it is possible to combine more than two base characters into one tie
bar, but this is not possible with Unicode (i.e. there is no possibility to draw
a tie bar over three of four characters). It is possible to indicate such larger
groups by repeating the tie bar, like in <a͡o͡u>. If really necessary, we
consider this possible, even though the rendering will never look good.

>>>>

   \item The uses of the apostrophe has led to long discussions on the Unicode
        Standard email list. An English keyboard inputs <{\fontspec{Monaco}'}>
        \textsc{apostrophe} at \uni{0027}, although the preferred Unicode
        apostrophe is the <\ ' > \textsc{right single quotation mark} at
        \uni{2019}.\footnote{Note that many word processors (like Microsoft
        Word) by default will replace straight quotes by curly quotes,
        depending on the whitespace around it.} However, the glottal
        stop/glottalization/ejective marker is yet another completely different
        character <{\large ʼ}>, the \textsc{modifier letter apostrophe} at
        \uni{02BC}, which unfortunately looks mostly extremely similar to
        \uni{2019}. 
  \item Another problem is the <ˁ> \textsc{modifier letter reversed glottal
        stop} at \uni{02C1} vs.\@ the <ˤ> \textsc{modifier letter small reversed
        glottal stop} at \uni{02E4}. Both appear in various resources
        representing phonetic data online. This is thus a clear example for
        which the Unicode Standard does not solve the linguistic standardization
        problem.
  \item Even linguists are unlikely to distinguish between the <ə>
        \textsc{latin small letter schwa} at code point \uni{0259} and <ǝ>
        \textsc{latin small letter turned e} at \uni{01DD}.
  \item The alveolar click <ǃ> at \uni{01C3} is of course often simply
        typed as <!> \textsc{exclamation mark} at \uni{0021}.\footnote{In the
        Unicode Standard the <ǃ> at \unif{01C3} is labeled \textsc{latin letter
        retroflex click}, but in IPA that glyph is used for an alveolar or
        postalveolar click (not retroflex). This naming is probably best seen as
        an error in the Unicode Standard.}
  \item The dental click <ǀ>, in Unicode known as \textsc{latin letter dental
        click} at \uni{01C0}, is of course often simply typed as <|>
        \textsc{vertical line} at \uni{007C}.
  \item For the marking of length there is a special Unicode characters, namely
        <ː> \textsc{modifier letter triangular colon} at \uni{02D0}. However,
        typing <:> \textsc{colon} at \uni{003A} is of course much easier.

@@

\begin{itemize}
  \item[] {\fontspec{Times New Roman}Times new Roman: <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] {\small \fontspec{CharisSIL}CharisSIL:\@ <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] {\footnotesize \fontspec{Monaco}Monaco: <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] {\fontspec{DoulosSIL}DoulosSIL:\@ <t̥ʰ͡s> <t̥ʰ͜s>}
  \item[] Linux Libertine: <t̥ʰ͡s> <t̥ʰ͜s>
\end{itemize}

Tie bars are a special type of character in the sense that they do not belong to
a segment, but bind two graphemes together. This actually turns out to be rather
different from the Unicode conceptions, see
Section~\ref{pitfall-different-notions-of-diacritics}. According to the spirit
of the IPA, it is possible to combine more than two base characters into one tie
bar, but this is not possible with Unicode (i.e.~there is no possibility to draw
a tie bar over three of four characters). It is possible to indicate such larger
groups by repeating the tie bar, like in <a͡o͡u>. If really necessary, we
consider this possible, even though the rendering will never look good.
